Efficient Data Provenance Analysis for Concurrent Programs

### 2013-11-15 paper.tex
added transmission case.  talked about comparative debugging.  seven benchmarks now

We illustrate the debugging benefits of last writer slices by using our system
to debug five real-world programs studied previously in the debugging
literature.  Table~\ref{tab:bugs} provides an overview of this evaluation and
Section~\ref{sec:eval:debugging:cases} looks in detail at several cases to 
better illustrate how last writer slices helps with debugging.

@@

The table also summarizes our evaluation of last writer slicing's debugging
benefits.  We mirror the evaluation strategy of bad value origin
tracking~\cite{badapples}, as it is the only other low-overhead provenance
tracking mechanism we know of.  We evaluate our system using their three
evaluation criteria - (1) {\bf effectiveness}, (2) {\bf triviality}, and (3)
{\bf usefulness}.  Our technique is {\bf effective} if it tracks a correct last
writer slice for the value or values involved in causing the failure.  A failure is {\bf
trivial} to debug if its cause is made obvious by examining the core dump
alone, {\em i.e.}, a last writer slice is unnecessary.  A last writer slice is
{\bf useful} if it is effective and the thread and code point in the slice help
understand why the failure occurred or how to fix it.   We directly compare to
bad value origin tracking.    The {\bf O.T.} column in Table~\ref{tab:bugs}
shows  whether their analysis could help debug each bug, based on the
description of their technique.

@@

The results in Table~\ref{tab:bugs} show that for both atomicity and ordering
violation bugs, last writer slices are effective.  In all cases, we found that
the last writer slice revealed the provenance of the data involved in the
operation that caused the failure, so last writer slicing was effective.  We
found that none of the failures we looked at were trivially debuggable.  This
finding is consistent with the difficulty of debugging concurrency bugs -- the
core dump does not reveal the interaction between threads that led to a
failure, only the result of the interaction, which leads to the failure.

Last writer slices were obviously useful in all five cases.  In these cases,
the root cause of the bug becomes clear when last writer slices show the
inter-thread data flow led to a variable having a value that caused a failure.
Using that information, we expect it would be straightforward in these cases to
add synchronization to fix these bugs.  Note that without last writer slices,
understanding the responsible inter-thread data flow and fixing the bug would
be difficult.

%For the remaining case, {\tt apache2}, last writer slices would
%probably help with debugging, but it is less obvious because this bug is a
%complex, multi-variable bug.  We provide more detail on {\tt apache2} in
%Section~\ref{sec:eval:debugging:cases}.
Our evaluation shows that bad value origin tracking~\cite{badapples} is not
helpful in three out of five cases where last writer slices is useful.  
There are two cases where bad value origin tracking may help.  For a
use-after-delete bug in {\tt pbzip2}, bad value origin tracking helps.  For
this bug, when one thread deletes a variable, it makes that variable unusable
and bad value origin tracking would track that value.  When the failure occurs,
the programmer would then know why the value was deleted.

@@

We now provide detailed cases studies showing the debugging role of last writer
slices in several of the bugs we examined.  Space constraints prevent us from
providing details for all cases.

@@

@@

\subsection{Characterizing \ctraps}
\label{sec:char}
We characterized the instrumentation pruned by escape analysis.  We report our
results in Table~\ref{tab:char}.  The table shows the percent of
instrumentation sites pruned and the absolute number of sites in parenthesis.

@@

\paragraph{Debugging with Data Provenance Information}
The most related prior work looked at using various forms of provenance information for debugging.

@@

Thin Slicing~\cite{thinslicing}, is a slicing technique that is related to our
work.  Thin Slicing yields the set of statements that computed on or wrote
values that influenced a particular variable's value (the ``seed'').  Thin
Slicing is similar in that its slices provide provenance and it excludes
most control information like our work.  Thin Slicing is different in that it has
overheads too high for production and focused on minimizing the size 
of each slice compared to prior slicing techniques.

@@

\section{Conclusions}
This work observes that we can collect a useful form of data provenance
information with overheads low enough for use in deployed systems.

>>>>

We illustrate the debugging benefits of last writer slices by using our system
to debug several real-world programs studied previously in the debugging
literature.  Table~\ref{tab:bugs} provides an overview of this evaluation and
Section~\ref{sec:eval:debugging:cases} looks in detail at several cases to 
better illustrate how last writer slices helps with debugging.

@@

The table also summarizes our evaluation of last writer slicing's debugging
benefits.  We mirror the evaluation strategy of bad value origin
tracking~\cite{badapples}, as it is the only other low-overhead provenance
tracking mechanism we know of.  We evaluate our system using their three
evaluation criteria - (1) {\bf effectiveness}, (2) {\bf triviality}, and (3)
{\bf usefulness}.  Our technique is {\bf effective} if it tracks a correct last
writer slice for the value or values involved in causing the failure.  A failure is {\bf
trivial} to debug if its cause is made obvious by examining the core dump
alone, {\em i.e.}, a last writer slice is unnecessary.  A last writer slice is
{\bf useful} if it is effective and the thread and code point in the slice help
understand why the failure occurred or how to fix it.   We directly compare to
bad value origin tracking~\cite{badapples}.    The {\bf O.T.} column in Table~\ref{tab:bugs}
shows  whether their analysis could help debug each bug, based on the
description of their technique.

@@

The results in Table~\ref{tab:bugs} show that for both atomicity and ordering
violation bugs, last writer slices are effective.  In all cases, we found that
the last writer slice revealed the provenance of at least some data involved in
the failure, so last writer slicing was effective.  We found that none of the
failures we looked at were trivially debuggable.  This finding is consistent
with the difficulty of debugging concurrency bugs -- the core dump does not
reveal the interaction between threads that led to a failure, only the result
of the interaction, which leads to the failure.

Last writer slices were useful for helping understand how to fix the bug in
all but one case and useful at least for understanding the bug in the
remaining case, {\tt apache2}.  The main reason last writer slices are helpful
is that they reveal the connection between code running in different threads
that interacts to cause a failure.  Without last writer slices, programmers
have no easy way of connecting these otherwise disparate parts of a program.
The case studies in Section~\ref{sec:eval:debugging:cases} show in more detail
how last writer slices are useful for debugging.

%For the remaining case, {\tt apache2}, last writer slices would
%probably help with debugging, but it is less obvious because this bug is a
%complex, multi-variable bug.  We provide more detail on {\tt apache2} in
%Section~\ref{sec:eval:debugging:cases}.
Our evaluation shows that bad value origin tracking~\cite{badapples} is
definitely helpful in one case where last writer slices is useful and may help
in one other case.  For a use-after-delete bug in {\tt pbzip2}, bad value
origin tracking helps.  For this bug, when one thread deletes a variable, it
makes that variable unusable and bad value origin tracking would track that
value.  When the failure occurs, the programmer would then know why the value
was deleted.

@@

We now provide detailed cases studies showing the debugging role of last writer
slices in several of the bugs we examined.

%Space constraints prevent us from providing details for all cases.

@@

\paragraph{transmission} We used last writer slices to debug a
use-before-initialization bug in transmission-1.42.  We omit a full case study
due to space constraints, but we mention a few notable properties of this case.
First, like {\tt httrack}, we debug {\tt transmission} by comparing the last
writer slice from failing and non-failing executions.  Second, last writer
slices reveal all the code and data involved in this bug, making the bug and
its fix clear.   Third, bad value origin tracking does not work in this case
because in failing executions, the bad value has no origin and in non-failing
executions, the involved variable holds only usable values.

@@

\subsection{Characterizing \ctraps}
\label{sec:char}
We characterized the instrumentation pruned by escape analysis.  We report our
results in Table~\ref{tab:char}.  The table shows the percent of
calls pruned and the absolute number of sites in parenthesis.

@@

\paragraph{Debugging with Data Provenance Information}
Prior work has looked at various forms of data provenance for debugging.

@@

Thin Slicing~\cite{thinslicing}, is a slicing technique that is related to our
work.  Thin Slicing yields the set of statements that computed on or wrote
values that influenced a ``seed'' variable's value.  Thin
Slicing is similar to our work in that slices provide provenance and it excludes most
control-flow information.  Thin Slicing is different in that it has
overheads too high for production and focused on minimizing the size of each
slice compared to prior slicing techniques.

@@

\section{Conclusions}
Last writer slices are the first system to collect a general form of data provenance information for concurrent programs with overheads low enough for production use.
